Recovering from population extinction in the Animal Life Cycle Algorithm (ALCA)

### 2022-09-30 paper.tex
first rev
Intro

\begin{document}

\title*{Recovering from population extinction in the Animal Life Cycle Algorithm (ALCA)} 
\titlerunning{Animal Life Cycle Algorithm (ALCA)}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{J. C. Felix-Saul, Mario García Valdez}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{J. C. Felix-Saul \at Tijuana Institute of Technology, Tijuana, Mexico, \email{jose.felix201@tectijuana.edu.mx}
\and Mario García Valdez \at Tijuana Institute of Technology, Tijuana, Mexico, \email{mario@tectijuana.edu.mx}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

@@

    Biologically inspired algorithms have proven to be very effective when
    solving complex optimization problems
    \cite{castillo2019comparative,valdez2021swarm,acherjee2020ultrasonic}, but
    the more challenging the problem, the more computing power it will require
    to solve them \cite{ontiveros2018high}. One technique to manage this
    processing power is to use distributed computing
    \cite{thain2005distributed} or the cloud available resources to help solve
    the problem \cite{garcia2013there,eshratifar2019bottlenet}. This strategy
    allows adaptating the required computing power according to the problem's
    complexity.

    Most frequently, bio-inspired algorithms are developed with a sequential
    (synchronous) perspective
    \cite{porto2018evolutionary,back1996evolutionary}, where a process must
    pause for the previous job to finish before continuing. In this research,
    we present an algorithm built as a native cloud solution, totally
    distributed, where its processes are executed asynchronously and in
    parallel. We designed the algorithm using the cloud available resources to
    distribute the processing workload among several computers or as a service
    in the cloud. This strategy allows to elastically increase (or reduce) the
    computing power according to the nature of the challenge
    \cite{armbrust2010view}.

    In previous work, we introduced an algorithm inspired by the biological
    life-cycle of animal species. As in nature, population individuals grow older
    and age. In this algorithm's reproduction process, couples must match by mutual
    attraction, where sometimes individuals won't procreate offspring because of
    the mate's low appeal. As time passes, the environment kills its individuals
    either because of low fitness or age, causing occasional population extinction.
    When this condition presents itself, it is required to create a new population
    to continue evolution (restart). We previously used to generate a new random
    set of solutions, with the cost of losing all evolution knowledge.

    The main contribution of this paper is to demonstrate that it is possible to
    evolve a population of individuals, similar to a Genetic Algorithm (GA), using
    a distributed, parallel, and asynchronous methodology by algorithm
    implementation and testing. We improve the algorithm behavior by solving
    population extinction caused by nature's pressure in the Animal Life Cycle
    Algorithm with one of the most promising alternatives: the projection of the
    Elite towards the Champion, where we compare the results obtained with the
    latter and the random use of the more traditional alternatives (mutation and
    crossover), using classic benchmark functions for optimization for comparison.

    We present the description of the following sections in our paper. First, we
    illustrate our algorithm model, the encountered problem with the occasional
    population extinction, and our proposed solution in
    section~\ref{section.proposal}, followed by our experimental configuration and
    results in section~\ref{section.experiments}, where we continue to analyze and
    describe some of our research findings in section~\ref{section.discussion}. We
    finalize by presenting some inferences based on the results of our experiments
    in section~\ref{section.conclusions}.

@@

% ---- Bibliography ----
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{bib/bibliografia.bib}

>>>>

\begin{document}
\title{Recovering from population extinction in the Animal Life Cycle Algorithm (ALCA)}
%

\titlerunning{Animal Life Cycle Algorithm (ALCA)}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{J. C. Felix-Saul, Mario Garcia Valdez}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{J. C. Felix-Saul \at Tijuana Institute of Technology, Tijuana, Mexico, \email{jose.felix201@tectijuana.edu.mx}
\and Mario García Valdez \at Tijuana Institute of Technology, Tijuana, Mexico, \email{mario@tectijuana.edu.mx}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

@@

    Biologically inspired algorithms have proven to be very effective when
    solving complex optimization problems
    \cite{castillo2019comparative,valdez2021swarm,acherjee2020ultrasonic}, but
    the more challenging the problem, the more computing power it will require
    to solve them \cite{ontiveros2018high}. % You can add a comment on why it  
    % requires more power, the main reason is the fitness evaluation function - M 
    One technique to manage this
    processing power is to use distributed computing % To manage the processing power need or demand? - M
    \cite{thain2005distributed} or the cloud resources 
    \cite{garcia2013there,eshratifar2019bottlenet}. This strategy
    allows adaptating the required computing power according to the problem's
    complexity.

    Traditionally, bio-inspired algorithms are developed with a sequential
    (synchronous) perspective
    \cite{porto2018evolutionary,back1996evolutionary}, where a process must
    pause for the previous task to finish before continuing.
    % Add here a few current works that use the cloud and other distributed technologies - M
    In this research,
    we present an algorithm built as a native cloud solution, totally
    distributed, where its processes are executed asynchronously and in
    parallel. We designed the algorithm using the cloud available resources to
    distribute the processing workload among several computers or as a service
    in the cloud. This strategy allows to elastically increase (or reduce) the
    computing power according to the nature of the challenge
    \cite{armbrust2010view}.

    In previous work, we introduced an algorithm inspired by the biological % Add the reference - M
    life-cycle of animal species. As in nature, population individuals grow older
    and age. In this algorithm's reproduction process, couples must match by mutual
    attraction, where sometimes individuals won't procreate offspring because of
    the mate's low appeal. As time passes, the environment kills its individuals
    either because of low fitness or age, causing occasional population extinction.
    % Just to be sure explain a bit more what we mean by extinsion. - M
    When this condition presents itself, it is required to create a new population
    to continue evolution (restart). 
    % Maybe you do it later, but explain what is a restart precisely 
    We previously used to generate a new random % In a previous work (reference) - M
    set of solutions, with the cost of losing all evolution knowledge.

    The main contribution of this paper is to demonstrate that it is possible to
    evolve a population of individuals, similar to a Genetic Algorithm (GA), using
    a distributed, parallel, and asynchronous methodology by algorithm
    implementation and testing.
    % The main contribution is not this, is the restarting method? - M 
    We improve the algorithm behavior by solving
    population extinction caused by nature's pressure in the Animal Life Cycle % This is more a contribution - M
    Algorithm with one of the most promising alternatives: the projection of the
    Elite towards the Champion, where we compare the results obtained with the
    latter and the random use of the more traditional alternatives (mutation and
    crossover), using classic benchmark functions for optimization for comparison.

    We present the description of the following sections in our paper. First, we % This paper is organized as follows - M

    illustrate our algorithm model, the encountered problem with the occasional
    population extinction, and our proposed solution in
    section~\ref{section.proposal}, followed by our experimental configuration and
    results in section~\ref{section.experiments}, where we continue to analyze and
    describe some of our research findings in section~\ref{section.discussion}. We
    finalize by presenting some inferences based on the results of our experiments
    in section~\ref{section.conclusions}.

@@

% ---- Bibliography ----
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%

\bibliographystyle{splncs03_unsrt}
%\bibliographystyle{splncs04}
\bibliography{bib/bibliografia.bib}
